Toward nonlinear multigrid \\ for nonlinear variational inequalities

### 2023-02-05 talk/slides.tex
talk progress finally

\section{variational inequalities (VIs) and obstacle problems}

\begin{frame}{the classical obstacle problem}

\begin{center}
\includegraphics[width=0.6\textwidth]{images/obstacle65.pdf}
\end{center}

\begin{frame}{active sets, free boundaries}

\begin{columns}
\begin{column}{0.6\textwidth}
\begin{itemize}
\item x
\end{itemize}
\end{column}
\begin{column}{0.4\textwidth}
\includegraphics[width=0.6\textwidth]{images/obstacle-sets.png}
\end{column}
\end{columns}
\end{frame}

@@

@@

\section{FAS multigrid for PDEs}

@@

\begin{itemize}
\item direct attack on the problem seems to require a \alert{multilevel} solver for \alert{variational inequalities} (VIs), but in the \alert{non-local residual case}
    \begin{itemize}
    \item[$\circ$] this seems not to exist
    \item[$\circ$] the \alert{smoother} must compute an implicit-step SKE residual including surface-motion term $\Phi(s) = - \bu|_s\cdot \bn_s$ from a scalable Stokes solver
    \end{itemize}
\item near-optimal multilevel VI solvers exist for toy problems
    \begin{itemize}
    \item[$\circ$] Poisson equation obstacle problem (Brandt \& Cryer 1983; Gr\"aeser \& Kornhuber 2009)
    \end{itemize}
\item<2> \alert{help me}
\end{itemize}
\end{frame}

>>>>

\section{variational inequalities (VIs)?}

\begin{frame}{example: a classical obstacle problem}

\includegraphics[width=0.6\textwidth]{images/obstacle65.pdf} \qquad \includegraphics[width=0.3\textwidth]{images/obstacle-sets.png}

\bigskip
\only<1>{
\begin{itemize}
\item \emph{problem.} on a domain $\Omega \subset \RR^2$, find the displacement $u(x)$ of a membrane, with fixed value $u = g$ on $\partial \Omega$, above an \emph{obstacle} $\psi(x)$, which minimizes the elastic energy
    $$J(u) = \int_\Omega \frac{1}{2} |\grad u|^2 - f\, u$$
\item shown above: $\Omega=(-2,2)^2$, $\psi(x)$ a hemisphere
\end{itemize}
}
\only<2>{
\begin{itemize}
\item i.e.~constrained optimization over an \emph{admissible set}
	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
\item informally, $J'(u)$ points directly into $\mathcal{K}$ (\emph{variational inequality}; VI):
    $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0 \quad \text{for all } v \in \mathcal{K}$$
\end{itemize}
}
\only<3>{
\begin{itemize}
\item the solution defines \emph{active} $A_u = \{u = \psi\}$ and \emph{inactive} $R_u = \{u> \psi\}$ subsets of $\Omega$, and a \emph{free boundary} $\Gamma_u=\partial R_u \cap \Omega$
\item a naive strong form poses the problem in terms of its solution:
{\small
\begin{align*}
-\grad^2 u &= f \text{ on $R_u$} \\
u &= \psi \text{ on $A_u$}
\end{align*}
}
\item by contrast, the weak form VI is meaningful
\end{itemize}
}
\only<4>{
\begin{itemize}
\item a \emph{complementarity problem}, essentially KKT conditions, is a meaningful strong form:
{\small
\begin{align*}
u - \psi &\ge 0 \\
-\grad^2 u - f &\ge 0 \\
(u - \psi)(-\grad^2 u - f) &= 0
\end{align*}
}
\end{itemize}
}
\end{frame}

\begin{frame}{general VI problems}

\begin{itemize}
\item let $\mathcal{K}$ be a closed and convex subset of a Banach space $\mathcal{V}$
\item suppose $F:\mathcal{K} \to \mathcal{V}'$ is a continuous, generally nonlinear \emph{operator}
    \begin{itemize}
    \item[$\circ$] $F$ may be defined only on $\mathcal{K}$
    \end{itemize}
\item the general problem VI($F$,$\mathcal{K}$) is
	$$\ip{F(u)}{v-u} \ge 0 \text{ for all } v \in \mathcal{K}$$

\bigskip
\item analogy:

\begin{tabular}{l|l}
x & y \\ \hline
z & w
\end{tabular}
\end{itemize}
\end{frame}

@@

\section{full approximation scheme (FAS) multigrid for PDEs?}

\begin{frame}{Bratu model problem: optimality}

\begin{columns}
\begin{column}{0.6\textwidth}
\begin{itemize}
\item x
\end{itemize}
\end{column}
\begin{column}{0.4\textwidth}
\includegraphics[width=\textwidth]{images/bratu-exps.png}

\includegraphics[width=\textwidth]{images/bratu-time.png}
\end{column}
\end{columns}
\end{frame}

@@

@@

\begin{itemize}
\item direct attack on the problem seems to require a \alert{multilevel} solver for \alert{variational inequalities} (VIs), but in the \alert{non-local residual case}
    \begin{itemize}
    \item[$\circ$] this seems not to exist
    \item[$\circ$] the \alert{smoother} must compute an implicit-step SKE residual including surface-motion term $\Phi(s) = - \bu|_s\cdot \bn_s$ from a scalable Stokes solver
    \end{itemize}
\item near-optimal multilevel VI solvers exist for toy problems
    \begin{itemize}
    \item[$\circ$] Poisson equation obstacle problem (Brandt \& Cryer 1983; Gr\"aeser \& Kornhuber 2009)
    \end{itemize}
\end{itemize}
\end{frame}
